---
title: "Alternate Fueling Stations Dataset - INITIAL TEMPLATE"
author: "M. Zhang"
date: "2022-03-01"
output:
  html_document:
    df_print: tibble
    toc: yes
  pdf_document:
    toc: yes
---

```{r, setup, include=FALSE}
library(tidyverse)
library(colorspace)
library(ggforce)
library(ggridges)
knitr::opts_chunk$set(echo = TRUE, comment = "")
```

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

# Introduction

We will be working with the `Alternative_Fueling_Stations.zip` dataset via the Tidy Tuesday repository. This analysis will not be styled as a "final report", but rather as a quick walk through of some data analysis and wrangling that took place while exploring the data.

The MSDSO Discord group for the University of Texas Masters in Data Science Online program will be doing weekly explorations of TidyTuesday as an exercise for improving their data science skill sets in a collaborative environment.

## Data Details

This dataset is pulled from the Tidy Tuesday Repository:

Thomas Mock (2022). Tidy Tuesday: A weekly data project aimed at the R ecosystem. https://github.com/rfordatascience/tidytuesday.

The original data is from [US DOT](https://data-usdot.opendata.arcgis.com/datasets/usdot::alternative-fueling-stations/about) and the [United Nations](https://unstats.un.org/unsd/methodology/m49/overview/) and has an accompanying [GIS map]](https://www.arcgis.com/home/webmap/viewer.html?panel=gallery&layers=cc51698ab9d94d67b4ec5dc5b8d97f34) and [EIA article](https://www.eia.gov/todayinenergy/detail.php?id=6050#:~:text=In%20aggregate%2C%20there%20are%20currently,gasoline%20stations%20in%20the%20country.).


## Data Loading

```{r Load, cache = TRUE}

tt <- tt_load("2021-11-23")

```


```{r, load-data, message = FALSE}
stations <- tt$stations
```


## Data Dictionary

See full LARGE data dictionary at: https://afdc.energy.gov/data_download/alt_fuel_stations_format

# Data Exploration and wrangling

## Examining the raw data

We will initially take a precursor inspection of the data, utilizing `summary` for numerical information and `table` for categorical information.

```{r, cursory-inspection}
glimpse(stations)
names(stations)
```

## Checking for NA values

```{r, examine_na}
stations %>%
  summarize(across(everything(), ~ sum(is.na(.)))) %>%
  tidyr::pivot_longer(everything()) %>%
  arrange(desc(value)) %>% 
  deframe()
```

NA counting methodology was taken from [stackexchange](https://stackoverflow.com/questions/63198917/using-tidyverse-pipeline-to-count-nas-and-reorder)


## Recoding

Going forward, we will be using the `Region_Name` column in lieu of `Region_Code` for simplicity. Additionally, we will recode `is_ldc` into categorical values. 

```{r, recode}

```

## Grouping candidates

Let's examine the total number of unique entries per column to see good candidates for faceting or other categorization methods.

```{r}
stations %>%
  sapply(n_distinct)
```

We can see that `LDC`, `Status` and `Region_Name` are all potential ways to cluster data.


## Developing Questions

While examining the data set, the following potential questions arose for investigation:

1.
2. 
3. 


# Data Visualizations

Pairwise plot for data - check raw data
```{r, pairplots-1, fig.width = 10, fig.height = 8}

```

Check grouped data 

```{r, pairplots-2, fig.width = 10, fig.height = 8}

```



# Conclusion




# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png")

```



