---
title: "Cryptic Crosswords in R (2022-04-19)"
date: 2022-05-06
output: html_document
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.

The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

This analysis was done as a follow along with David Robinson's screencast of Tidy Tuesday (https://youtu.be/T11WKngDEGQ)

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(tidytext)
library(tidylo)
library(clock)

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tuesdata <- tidytuesdayR::tt_load('2022-04-19')

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tuesdata 

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tuesdata %>% 
  map(glimpse)

```

# Wrangle

For this analysis we will parse the lengths (word and character) for clues, definitions, and answers.

```{r Wrangle}

times <- tuesdata$times
big_dave <- tuesdata$big_dave

# clean data and calculate word lengths
crosswords <- times %>%
  bind_rows(big_dave) %>% 
  select(-rowid) %>%
  extract(clue, c("clue", "answer_length"), "(.*) \\((.*)\\)$") %>%
  mutate(year = get_year(puzzle_date),
         month = get_month(puzzle_date),
         month_date = date_group(puzzle_date, "month"),
         definition_idx = str_locate(clue, fixed(definition))[, 1],
         definition_position = case_when(
           definition_idx == 1 ~ 0,
           is.na(definition_idx) ~ NA_real_,
           TRUE ~ 1),
         definition_length = str_length(definition),
         clue_length = str_length(clue),
         n_words_answer = lengths(str_split(answer, " ")),
         n_words_definition = lengths(str_split(definition, " ")),
         n_words_clue = str_count(clue, " ") + 1
         )

crosswords %>% 
  select(clue, n_words_clue, answer_length, n_words_answer, answer, n_words_definition, definition) %>%
         head()

```


# Visualize

Examining the overall data source distribution and looking into missing data

```{r Visualize}
# visualize number of puzzles
crosswords %>% 
  mutate(date = date_group(puzzle_date, "month", n = 3)) %>%
  group_by(date) %>%
  count(source) %>%
  ggplot() + 
  geom_col() + 
  aes(x = date, y = n, fill = source) + 
  ggtitle("number of entries by quarter")

crosswords %>%
  group_by(year, source) %>%
  summarize(pct_missing = mean(is.na(definition_position))) %>%
  ggplot() +
  aes(x = year, y = pct_missing, color = source) + 
  geom_line() + 
  ggtitle("percent of puzzles missing definitions")

```

More puzzles are provided by "big_dave" and more recent data has less missing data than earlier data.


Where do definitions typically position themselves within clues?

```{r}
crosswords %>%
  ggplot() + 
  aes(x = definition_position, fill = source) +
  geom_histogram(center = 0.5, position = "dodge") + 
  ggtitle("Definition at beginning (0) or end (1) of clue")

crosswords %>%
  summarize(definition_distribution = mean(definition_position, na.rm = TRUE))

crosswords %>%
  group_by(year, source) %>%
  summarize(definition_distribution = mean(definition_position, na.rm = TRUE)) %>%
  ggplot() + 
  geom_line() +
  aes(x = year, y = definition_distribution, color = source) + 
  ggtitle("Average position of definition by year")

crosswords %>%
  group_by(year, source) %>%
  filter(source == "bigdave44") %>%
  summarize(definition_distribution = mean(definition_position, na.rm = TRUE)) %>%
  glm(definition_distribution ~ year, data = .) %>%
  summary()
```

Positioning of definition within clues over years is slowly trending to favor being at the end of the clue.


```{r}
crosswords %>%
  ggplot() +
  geom_histogram() +
  aes(definition_length) + 
  facet_wrap(vars(source), nrow = 2) +
  ggtitle("distribution of definition character length")

crosswords %>%
  filter(!is.na(definition)) %>%
  group_by(source) %>%
  summarize(median_length = median(definition_length))
```


Is there a relationship between clue length and definition length?

Is there a relationship with the number of words in the definition and the number of words in the clue?

```{r}
crosswords %>%
  ggplot() + 
  aes(x = clue_length, y = definition_length) + 
  geom_density_2d_filled() + 
  geom_density_2d() + 
  scale_x_continuous(limits = c(0, 75)) +
  scale_y_continuous(limits = c(0, 75)) + 
  ggtitle("relationship between definition and clue character length")

crosswords %>%
  with(cor.test(definition_length, clue_length))

crosswords %>%
  with(cor.test(n_words_definition, n_words_clue))

crosswords %>%
  with(cor.test(n_words_answer, n_words_definition))

```

There is a weak correlation between clue length and definition length, for both characters and word count.


What are the most common words used as answers in cryptic crosswords?

```{r}
answer_words <- crosswords %>%
  filter(!is.na(answer)) %>%
  unnest_tokens(output=word, input=answer)

answer_words %>%
  count(source, word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  group_by(word) %>%
  summarize(source, n, n_total = sum(n)) %>%
  ungroup() %>%
  top_n(20, n_total) %>%
  ggplot() +
  aes(x = n, y = fct_reorder(word, n_total), fill = source) + 
  geom_col() + 
  ggtitle("Most Common Words in Answer")

```

Are the words used for answers different between sets?

```{r}

answer_words %>%
  count(source, word, sort = TRUE) %>%
  anti_join(stop_words, by = "word") %>%
  group_by(word) %>%
  summarize(source, n, n_total = sum(n)) %>%
  group_by(source) %>%
  slice_max(n = 10, order_by = n) %>%
  ggplot() +
  aes(x = n, y = fct_reorder(word, n), fill = source) + 
  geom_col() + 
  facet_wrap(vars(source), scale = "free") + 
  ggtitle("Most common words in Answer by Source")

answers_log_odds <- answer_words %>%
  count(source, word) %>%
  anti_join(stop_words, by = "word") %>%
  bind_log_odds(source, word, n)

answers_log_odds %>%
  arrange(desc(abs(log_odds_weighted))) %>%
  group_by(source) %>%
  filter(n > 50) %>%
  top_n(10, log_odds_weighted) %>%
  ungroup() %>%
  mutate(word = fct_reorder(word, log_odds_weighted)) %>%
  ggplot() + 
  aes(log_odds_weighted, word) +
  geom_col() +
  facet_wrap(vars(source), scales = "free") +
  ggtitle("Words favored in one source which appear over 50x")

```

