---
title: "Eurovision Dataset Analysis"
date: 2022-05-17
output: html_document
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(colorspace)

```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load}

tt <- tt_load("2022-05-17")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}

tt %>% 
  map(glimpse);

```

# Wrangle and Visualize


```{r Wrangle}
eurovision <- tt$eurovision
votes <- tt$"eurovision-votes"

```

What is the trend of number of songs submitted per year?

```{r}
# colorspace::choose_color()
colors <- c('#629F55', '#6394C9', '#CD7792')

eurovision_year <- eurovision %>%
  group_by(year, artist_country) %>%
  mutate(song_count = n()) %>% 
  ungroup() %>%
  group_by(year) %>%
  summarize(songs_count = n(),
            country_count = n_distinct(artist_country),
            artist_count = n_distinct(artist)) %>%
  pivot_longer(cols = c(songs_count, country_count, artist_count), names_to = "count_type", values_to = "count") 

eurovision_year

eurovision_year %>%
  ggplot() + 
  geom_line(size = 1.5, alpha = 0.6) + 
  aes(x = year, y = count, color = count_type) + 
  theme_bw() + 
  scale_x_continuous(name = "Year") + 
  scale_y_continuous(name = "Count") + 
  scale_color_manual(values = colors, name = NULL, labels = c("Participating artists", "Participating countries", "Songs submitted")) + 
  ggtitle("Eurovision participation over time")

eurovision %>%
  group_by()

eurovision %>%
  group_by(artist_country) %>%
  summarize(n = n())

```

How often did countries participate in Eurovision and how did they perform?

```{r}
# map the host countries

library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(ggiraph)

```

```{r}
map_countries <- ne_countries(scale = "medium", continent = c("asia", "europe"), returnclass = "sf")


hostcountry_eurovision <- eurovision %>% 
  group_by(host_country) %>%
  summarize(n_hosted = n_distinct(year))
  
artistcountry_eurovision <- 
  eurovision %>% 
  group_by(artist_country) %>%
  mutate(n_year = n_distinct(year)) %>%
  ungroup() %>%
  group_by(artist_country, section) %>%
  summarize(first = sum(rank == 1, na.rm = TRUE),
            second = sum(rank == 2, na.rm = TRUE),
            third = sum(rank == 3, na.rm = TRUE),
            grand_champion = sum(winner),
            n_participated = n(),
            n_year = first(n_year)) %>%
  ungroup() %>%
  group_by(artist_country) %>%
  summarize(n_participated = sum(n_participated),
            first = sum(first),
            second = sum(second),
            third = sum(third),
            grand_champion = sum(grand_champion),
            n_year_participated = first(n_year)) %>%
  mutate(top3 = first + second + third,
         percent_top3 = top3/n_participated)


eurovision_summary <- hostcountry_eurovision %>%
  full_join(artistcountry_eurovision, by = c("host_country" = "artist_country")) %>%
  rename(country = host_country) %>%
  mutate(n_hosted = replace_na(n_hosted, 0))

# eurovision_summary %>% View()


map_eurovision <- eurovision_summary %>%
  right_join(map_countries %>% select(admin, geometry), by = c("country" = "admin")) %>%
  filter(!country %in% c("Antarctica")) %>%
  st_as_sf(sf_column_name = c("geometry")) %>%
  mutate(
         tooltip = case_when(
           !is.na(n_participated) ~ paste0(country, "\n",
                                           "Participated: ", n_participated, "\n",
                                           "Years: ", n_year_participated, "\n",
                                           "Hosted: ", n_hosted, "\n",
                                           "First: ", first, "\n",
                                           "Second: ", third, "\n",
                                           "Third: ", second, "\n",
                                           "Podium % (Top 3): ", signif(percent_top3, 3)*100, "%"
                                           ),
           TRUE ~ paste0(country, ": Non Participant")
           ),
         n_participated = replace_na(n_participated, 0)
  )

```

Interactive map: 

```{r}
map_eurovision_interactive <- map_eurovision %>%
  ggplot() + 
  geom_sf_interactive() + 
  aes(fill = n_participated, 
      tooltip = tooltip, 
      data_id = tooltip) +
  coord_sf(xlim = c(-40, 180), ylim = c(30, 90), expand = FALSE) + 
  scale_fill_continuous_sequential(palette = "Blues", name = "Times Participated") +
  ggtitle("Eurovision Participation",
          subtitle = "Interactive Map")

girafe(
  ggobj = map_eurovision_interactive, 
  width_svg = 9, height_svg = 9*0.618,
  options = list(
    opts_tooltip(css = "background: #F5F5F5; color: #191970")
    ))
```

```{r}
eurovision_summary %>%
  mutate(country = fct_reorder(country, n_participated)) %>%
  top_n(20, n_participated) %>%
  ggplot() + geom_col() + 
  aes(x = n_participated, y = country) + 
  xlab("Times participated") +
  ylab("Country")

eurovision_summary %>%
  mutate(country = fct_reorder(country, n_hosted)) %>%
  top_n(20, n_hosted) %>%
  ggplot() + geom_col() + 
  aes(x = n_hosted, y = country) + 
  xlab("Times Hosted") +
  ylab("Country")

eurovision_summary %>%
  mutate(country = fct_reorder(country, percent_top3)) %>%
  top_n(20, percent_top3) %>%
  ggplot() + geom_col() + 
  aes(x = percent_top3, y = country) + 
  xlab("Percentage of participants in Top 3") +
  ylab("Country") + 
  scale_x_continuous(labels = scales::percent)

```

