---
title: "TidyTuesday (2021-11-23) Dr. Who"
date: 2022-03-14
output: html_document
editor_options: 
  chunk_output_type: console
---

# TidyTuesday

Join the R4DS Online Learning Community in the weekly #TidyTuesday event!
Every week we post a raw dataset, a chart or article related to that dataset, and ask you to explore the data.
While the dataset will be “tamed”, it will not always be tidy! As such you might need to apply various R for Data Science techniques to wrangle the data into a true tidy format.
The goal of TidyTuesday is to apply your R skills, get feedback, explore other’s work, and connect with the greater #RStats community!
As such we encourage everyone of all skills to participate!

This was a follow along with [David Robinson's analysis](https://youtu.be/89l_3Y53qMM) on the data set.

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(tidytuesdayR)
library(ggforce)
library(ggridges)
theme_set(theme_classic())
```

# Load the weekly Data

Download the weekly data and make available in the `tt` object.

```{r Load, cache = TRUE}

tt <- tt_load("2021-11-23")

```


# Readme

Take a look at the readme for the weekly data to get insight on the dataset.
This includes a data dictionary, source, and a link to an article on the data.

```{r Readme, eval = interactive()}

tt

```


# Glimpse Data

Take an initial look at the format of the data available.

```{r Glimpse}
tt %>% 
  map(glimpse)

```

```{r}
directors <- tt$directors %>% janitor::clean_names()
episodes <- tt$episodes %>% janitor::clean_names() %>%
  fill(season_number)%>%
  mutate(episode_id = paste0(season_number,".",coalesce(as.character(episode_number), "X"),".",episode_title),
        episode_id = fct_reorder(episode_id, first_aired),
        episode_title = fct_reorder(episode_title, first_aired),
        overall_episode_number = as.integer(episode_title))
writers <- tt$writers %>% janitor::clean_names()
imdb <- tt$imdb %>% janitor::clean_names()
```

```{r}
episodes %>%
  select(first_aired, uk_viewers, rating, duration) %>%
  summary()

episodes %>% select(era) %>% table(useNA = "ifany")
episodes %>% select(season_number) %>% table(useNA = "ifany")
episodes %>% select(episode_number) %>% table(useNA = "ifany")
episodes %>% select(serial_title) %>% table(useNA = "ifany")
episodes %>% select(story_number) %>% table(useNA = "ifany")
episodes %>% select(type) %>% table(useNA = "ifany")
episodes %>% select(production_code) %>% table(useNA = "ifany")
# episodes %>% select(episode_id) %>% table(useNA = "ifany")

```

## Examining NA values


```{r}
episodes %>%
  filter(is.na(production_code)) %>% View()
```

By looking at information which was previously unavailable (episode_number, season_number, serial_title, production_code) we can see that when `type == special` there is no episode_number or season_number. The only populated values for `serial_title` are ` == "flux"` and are for the latest season 13, and that production_code was dropped during season 7.

We will be removing `serial_title`, `era`, and `production_code` since they will not be used in any future analysis for this set.

```{r}
episodes <- episodes %>%
  select(-serial_title, -production_code, -era)

episodes %>% names()
```

Then re-examine NA values:

```{r, examine_na}
episodes %>%
  summarize(across(everything(), ~ sum(is.na(.)))) %>%
  tidyr::pivot_longer(everything()) %>%
  arrange(desc(value)) %>% 
  deframe()
```

NA counting methodology was taken from [stackexchange](https://stackoverflow.com/questions/63198917/using-tidyverse-pipeline-to-count-nas-and-reorder)


# Visualize

Using your processed dataset, create your unique visualization.

We will examine the viewership across all the seasons.

```{r Visualize_viewers}
episodes %>%
  #filter(season_number <= 4) %>%
  ggplot(aes(episode_title, uk_viewers, fill = factor(season_number))) + 
  geom_col() + 
  scale_x_discrete(name = "") +
  scale_y_continuous(name = "UK Viewers (mln)") + 
  scale_fill_discrete(name = "Season") + 
  #theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  theme(axis.text.x = element_blank()) + 
  ggtitle("UK Viewers over time for Dr. Who")
```


Similarly, look at ratings:


```{r Visualize_rating}
episodes %>%
  ggplot(aes(as.numeric(episode_title), rating)) + 
  geom_point(size = 3, aes(color = factor(season_number))) + 
  geom_line(group = 1) + 
  geom_smooth(method = "loess") + 
  geom_text(aes(label = episode_title), hjust = 1, vjust = 1, check_overlap = TRUE) + 
  scale_x_discrete(name = "Episode") +
  scale_y_continuous(name = "Average Ratings") + 
  scale_color_discrete(name = "Season") + 
  theme(axis.text.x = element_blank()) + 
  ggtitle("Ratings over time for Dr. Who")
  

```

```{r}
summarize_episodes <- function(tbl) {
  tbl %>%
    summarize(median_rating = median(rating, na.rm = TRUE),
              average_rating = mean(rating, na.rm = TRUE),
              median_viewers = median(uk_viewers, na.rm = TRUE),
              average_viewers = mean(uk_viewers, na.rm = TRUE),
              n_episodes = n(),
              t_test = list(broom::tidy(t.test(rating[!is.na(rating)])))
              ) %>%
              unnest(t_test) %>%
  arrange(desc(n_episodes))
}
```

```{r}
episodes %>%
  inner_join(writers, by ="story_number") %>%
  group_by(writer = fct_lump(writer, 6)) %>%
             summarize_episodes() %>%
  mutate(writer = fct_reorder(writer, average_rating)) %>%
  ggplot() + 
  aes(x = average_rating, y = writer) + 
  geom_point() + 
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.1) + 
  scale_x_continuous(name = "Average Rating (w/ 95% confidence intervals)") +
  scale_y_discrete(name = "Writers") + 
  ggtitle("Ratings by Writer")
```

```{r}
episodes %>%
  inner_join(writers, by ="story_number") %>%
  mutate(writer = fct_lump(writer, 6),
         writer = fct_reorder(writer, as.integer(episode_title))) %>%
  filter(writer != "Other") %>%
  ggplot() + 
  aes(x = as.integer(episode_title), y = rating) + 
  geom_point(size = 2) + 
  geom_line(data = episodes, alpha = 0.3) + 
  facet_wrap(~ writer) + 
  scale_x_continuous(name = "") +
  scale_y_discrete(name = "Writers") + 
  ggtitle("Ratings by Writer by Episodes")
```

Statistical tests for whether writer makes unpopular or popular episodes?

```{r}
writers_spread <- episodes %>%
  inner_join(writers, by ="story_number") %>%
  mutate(writer = fct_lump(writer, 6),
         value = 1) %>%
  filter(writer != "Other") %>%
  pivot_wider(names_from = "writer", values_from = "value", values_fill = 0) %>%
  janitor::clean_names()
  #spread(writer, value, fill = 0) %>% View()
```

```{r}
episodes %>% 
  ggplot() +
  aes(overall_episode_number, rating) +
  geom_line() + 
  geom_smooth(method = "gam", formula = y ~ s(x, k = 3))

library(splines)

episode_predictions <- lm(rating ~ 
     ns(overall_episode_number, 3), episodes) %>%
  broom::augment(data = episodes %>% filter(!is.na(rating))) %>%
  select(episode_title, story_number, rating, .fitted, .resid)
     
episode_predictions %>%
  inner_join(writers) %>%
  mutate(writer = fct_lump(writer, 6),
         writer = fct_reorder(writer, .resid)) %>%
  ggplot(aes(.resid, writer)) + 
  geom_boxplot() + 
  scale_x_continuous(name = "Residual Rating Relative to Position in Series")
  
lm(rating ~ 
   ns(overall_episode_number, 3) + 
   russell_t_davies + 
   mark_gatiss + 
     chris_chibnall + 
     steven_moffat,
   writers_spread) %>%
  #aov() %>%
  summary()

```



# Save Image

Save your image for sharing. Be sure to use the `#TidyTuesday` hashtag in your post on twitter! 

```{r}

# This will save your most recent plot
ggsave(
  filename = "My TidyTuesday Plot.png",
  device = "png")

```
